---
title: "Tests for associations with pupil shape"
author: "Katie Thomas"
date: 21 January 2022
output:
 html_document:
    keep_md: true
    theme: flatly
    toc: yes
    toc_float: yes
---

<style type="text/css">

body{ /* Normal  */
      font-size: 17px;
  }
  
</style>

```{r setup, include = FALSE}

#global rmarkdown options
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

#load packages
library(ape)
library(geiger)
library(picante)
library(caper)
library(geiger)
library(phytools)
library(plotly)
library(knitr)
library(kableExtra)
library(gridExtra)
library(grid)
library(gtable)
library(cowplot)
library(plyr)
library(tidyverse)
library(grImport)
library(AmphiNom)
library(phylolm)
```

# Data

## Import cleaned data and tree

```{r}
#import cleaned data
pupil.data <- data.frame(read.csv("../Data/Cleaned data/pupil_data_refs.csv",header=TRUE, na.strings=c("NA")))

#import cleaned tree
pupil.tree <- read.nexus(file = "../Data/Cleaned data/pupil_tree_cleaned")
```


## Subset data

### Ecology

First we subset our data to include only species with ecological trait data (n = 909) and make a corresponding species tree. 

```{r}

# Subset data to species with ecological trait data -----
pupils.eco <- pupil.data %>%
  filter_at(vars(aquatic, diurnal, fossorial, arboreal), any_vars(!is.na(.)))  #omit species missing all adult ecology data.

# Prune tree to match data subset -------

#make row names of the datafram the phylogeny tip labels
rownames(pupils.eco) <- pupils.eco$ASW_names

#make list of taxa to drop (in tree but not in dataset)
drops <- setdiff(pupil.tree$tip.label, pupils.eco$ASW_names)

#drop unwanted tips from phylogeny
eco.tree <- drop.tip(phy = pupil.tree, tip = drops) 

#check that tree tip labels match data subset
name.check(eco.tree, pupils.eco)

#reorder data to match tip labels
pupils.eco <- pupils.eco[eco.tree$tip.label,]
```


### Eye size

Next we subset our data to species that have eye size data available from our previous work (n = 207; Thomas et al. 2020).

```{r eyesize-subset, results = "hide"}

# Subset data to species with eye size data only -----
pupils.eye <- pupil.data %>%
  filter(!is.na(eye_av)) %>%
  mutate(binary_constriction = recode(Final_Constriction,"horizontal" = "elongated", "vertical" = "elongated")) 

# Prune tree to match data subset -------

#make row names of the datafram the phylogeny tip labels
rownames(pupils.eye) <- pupils.eye$ASW_names

#make list of taxa to drop (in tree but not in dataset)
drops <- setdiff(pupil.tree$tip.label, pupils.eye$ASW_names)

#drop unwanted tips from phylogeny
eye.tree <- drop.tip(phy = pupil.tree, tip = drops) 

#check that tree tip labels match data subset
name.check(eye.tree, pupils.eye)

#reorder data to match tree
pupils.eye <- pupils.eye[eye.tree$tip.label,]
```


# Adult pupils across ecology

## Visualize data

Here, we plot adult pupil constriction  alongside binary ecological trait data for whether each species is s, aquatic, fossorial, or diurnal. Species names are colored by pupil constriction and each binary trait is  colored as present (colored) or absent (gray). Missing trait data is left as white space.

```{r}
# Designate color and shape vectors for plotting -----

#pupil constriction colors
col_constrict <- c("horizontal" = "#f768fc",
                   "symmetrical" = "#ffba15",
                   "vertical" = "#3abde2")

#pupil constriction symbols
sh_constrict <- c("horizontal" = "H",
                  "symmetrical" = "N",
                  "vertical" = "V")
                   
#pupil shape symbols
sh_shape <- c("almond/slit" = 97,
              "circle"  = 19, 
              "diamond" = 23, 
              "sideways triangle" = 24,
              "upside down tear" = 6,
              "upside down triangle" = 25)

#aquatic cols
col_aq <- c("no" = "gray50",
            "yes" = "#0072B2")

#fossorial cols
col_foss <- c("no" = "gray50",
              "yes" = "#D55E00")

#scansorial cols
col_scans <- c("no" = "gray50",
               "yes" = "#009E73")

#diurnal cols
col_diur <- c("no" = "gray50",
              "yes" = "#FFAF27")

```


```{r, fig.height=17, fig.width=17}

#export as pdf
#pdf("../Outputs/Figures/ecology_phylo.pdf", width = 20, height = 20)

#plot phylogeny
plot.phylo(eco.tree, 
           type = "fan",
           show.tip.label = TRUE, 
           cex = 0.4,
           tip.color = col_constrict[pupils.eco$Final_Constriction], #
           use.edge.length = TRUE, 
           edge.width = 0.3,
           label.offset = 30) 

#add tip labels for aquatic
tiplabels(col = col_aq[pupils.eco$aquatic],
          pch = 19, 
          cex = 0.6,
          offset = 5) 

#add tip labels for fossorial
tiplabels(col = col_foss[pupils.eco$fossorial],
          pch = 19, 
          cex = 0.6,
          offset = 10) 

#add tip labels for scansorial
tiplabels(col = col_scans[pupils.eco$arboreal],
          pch = 19, 
          cex = 0.6,
          offset = 15) 

#add tip labels for diurnal
tiplabels(col = col_diur[pupils.eco$diurnal],
          pch = 19, 
          cex = 0.6,
          offset = 20) 

#add legend for pupil constriction
legend(x = 0.05, y = -20, legend = c("Horizontally elongated", "Non-elongated", "Vertically elongated"), 
       col = col_constrict,
       pch = 15,
       cex = 0.7, 
       box.lty = 0, 
       title = "Pupil constriction", 
       title.adj = 0)

#add legend for ecology
legend(x = 0.05, y = -50, legend = c("aquatic", "fossorial", "scansorial", "diurnal", "absent"), 
       col = c("#0072B2","#D55E00","#009E73", "#FFAF27","gray50"),
       pch = 19,
       cex = 0.7, 
       box.lty = 0, 
       title = "Adult Ecology", 
       title.adj = 0)

#finish pdf export
#dev.off()
```

## Adult pupil shape and ecology

We use multivariate phylogenetic logistic regression using the R package phylolm to fit models of binary pupil shape vs. discrete categories for ecology. We use the logistic_MPLE method, which maximizes the penalized likelihood of the logistic regression.

Our ecological data is already in binary format, with "yes" indicating the presence of an ecological trait of interest and "no" the absence. We also need to make pupil shape binary depending on the hypothesis we are testing. Specifically, we need one column for vertical pupil vs. nonvertical (horizontal or symmetrical) pupil, and one column for symmetrical pupil vs. nonsymmetrical (elongated horizontal or vertical) pupil.

```{r}
#convert traits to 0 and 1
pupils.binary <- pupils.eco %>%
  select(ASW_names, Final_Constriction, aquatic, fossorial, arboreal, diurnal) %>%
  mutate(pupil_vert_bi = recode(Final_Constriction, "vertical" = 1, "horizontal" = 0, "symmetrical" = 0)) %>%
  mutate(pupil_symm_bi = recode(Final_Constriction, "vertical" = 0, "horizontal" = 0, "symmetrical" = 1)) %>%
  
  mutate(aquatic_bi = recode(aquatic, "yes" = 1, "no" = 0)) %>%
  mutate(fossorial_bi = recode(fossorial, "yes" = 1, "no" = 0)) %>%
  mutate(arboreal_bi = recode(arboreal, "yes" = 1, "no" = 0)) %>%
  mutate(diurnal_bi = recode(diurnal, "yes" = 1, "no" = 0)) %>%
  mutate(foss_aq_bi = if_else(fossorial=="yes"|aquatic=="yes", 1, 0)) %>%
  mutate(three_state_eco = case_when(fossorial=="yes"~"fossorial",
                                     aquatic=="yes"~"aquatic",
                                     fossorial=="no"&aquatic=="no"~"neither"))
```

Now we can run the phylogenetic GLMs. We will subset these by the hypothesis that is being tested.

## Are vertical pupils correlated with scansorial habitats?

First we take a quick look at the distribution of all 4 combinations of these 2 binary traits. 

```{r}

#find counts for each combination of states
counts_scans <- pupils.binary %>%
  filter(!is.na(arboreal_bi)) %>%
  mutate(dual_state = case_when(pupil_vert_bi==1 & arboreal_bi==1 ~ "vertical & scansorial",
                                pupil_vert_bi==0 & arboreal_bi==1 ~ "nonvertical & scansorial",
                                pupil_vert_bi==1 & arboreal_bi==0 ~ "vertical & nonscansorial",
                                pupil_vert_bi==0 & arboreal_bi==0 ~ "nonvertical & nonscansorial"))

#print table
kable(count(counts_scans, dual_state))

#plot states
plot_S <- ggplot(data=count(counts_scans, dual_state), 
                  aes(x=n, y=dual_state)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=n), hjust=0)+
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black"), legend.key = element_rect(fill = NA)) + #controls background +
  xlab("Number of species") +
  ylab("Pupil constriction and habitat")

plot(plot_S)

```

Then we can run the model in phyloglm. For this model, pupil shape is binary (1 = vertical pupils, 0 = horizontal or non-elongated pupils) and scansoriality is binary (1 = scansorial, 0 = non-scansorial)

```{r}
#fit model
ScansGLM <- phyloglm(pupil_vert_bi ~ arboreal_bi, 
                       data = pupils.binary, 
                       phy = eco.tree, 
                       method = "logistic_MPLE", 
                       boot = 1000)
#output
summary(ScansGLM)
```


## Tests for correlates of non-elongated pupils

We predicted that non-elongated (e.g. circular) pupils would show evolutionary associations with 1) aquatic habitats, 2) fossorial habits, and 3) diurnal activity periods. 

The below models have a binary, discrete category for pupil shape (the dependent variable; 1 = non-elongated pupil, 0 = elongated [horizontal or vertical] pupil). Covariates are 1) habitat, a discrete variable with 3 states (aquatic, fossorial, neither), and 2) activity period, a discrete variable with 2 states (1 = diurnal, 0 = non-diurnal)

Here, we test three models of the evolution of binary pupil shape (non-elongated or elongated):

1) pupil shape ~ habitat
2) pupil shape ~ activity period
3) pupil shape ~ habitat + activity period

Preliminary fits for the second and third models listed prompted warning messages that ""the estimated coefficients in the absence of phylogenetic signal lead to some linear predictors beyond 'btol'." Thus, we increased btol (the bound on the linear predictor to bound the searching space) to 25 (default is btol = 10) for all three models. 

```{r}
#control factor level order for eco states
pupils.binary$three_state_eco <- factor(pupils.binary$three_state_eco, levels = c("neither", "fossorial", "aquatic"))

#subset data to species with complete sampling for ecology and diurnality
pupils.binary_habact <- pupils.binary %>%
  filter(across(c(three_state_eco, diurnal_bi), ~ !is.na(.x)))

#sample sizes for each combination of states
samples <- pupils.binary_habact %>%
  count(three_state_eco, diurnal_bi)

#make scrolling table
kable(samples, caption = "Number of obsercations in each combination of states for model. ") %>%
  kable_styling(full_width = F) %>%
  collapse_rows(columns = 1, valign = "top") %>%
  scroll_box(height = "500px")
```


### Pupil ~ habitat

```{r}

#run model
ecoGLM_hab <- phyloglm(pupil_symm_bi ~ three_state_eco, 
                       data = pupils.binary_habact, 
                       phy = eco.tree, 
                       btol = 25,
                       method = "logistic_MPLE", 
                       boot = 1000)
#output
summary(ecoGLM_hab)
```

### Pupil ~ activity period

```{r}

#run model
ecoGLM_act <- phyloglm(pupil_symm_bi ~ diurnal_bi, 
                       data = pupils.binary_habact, 
                       phy = eco.tree, 
                       btol = 25,
                       method = "logistic_MPLE", 
                       boot = 1000)

#output
summary(ecoGLM_act)
```

### Pupil ~ habitat + activity period

```{r}

#run model
ecoGLM_habact <- phyloglm(pupil_symm_bi ~ three_state_eco + diurnal_bi, 
                       data = pupils.binary_habact, 
                       phy = eco.tree, 
                       btol = 25,
                       method = "logistic_MPLE", 
                       boot = 1000)
summary(ecoGLM_habact)
coef(ecoGLM_habact)
vcov(ecoGLM_habact)
```

### Model comparison

Then we can compare the two models via AIC and log likelihood. 

```{r}

#model AIC comparison
aics <- c("habitat" = AIC(ecoGLM_hab),
          "activity per." = AIC(ecoGLM_act),
          "hab. + act. per." = AIC(ecoGLM_habact))

#print aics
aics
```

The model including both habitat and activity period is a significantly better fit than the one including only habitat, which outperforms the one including only activity period based on a significance criteria of delta AIC > 2.


# Adult pupil shape and eye size

We predicted that species with larger eyes would benefit from having a slit pupil to facilitate a large pupillary range. 

## Pupil shape and eye size

Here, we plot pupil shape and absolute eye size (scaled purple circles). 

```{r, fig.height=40, fig.width=7}
#export as pdf
#pdf("../Outputs/Figures/phylogeny_eyesize.pdf", width = 8, height = 30)

#plot phylogeny
plot.phylo(eye.tree, 
           type = "phylogram", 
           show.tip.label = TRUE, 
           cex = 0.8, #text size
           no.margin = TRUE, 
           use.edge.length = TRUE, 
           edge.width = 2,
           label.offset = 75) 

#add tip labels for pupil constriction
tiplabels(col = col_constrict[pupils.eye$Final_Constriction], #sets color to pupil constriction 
          pch = sh_constrict[pupils.eye$Final_Constriction], #shape of labels
          cex = 0.8,
          offset = 5) #size of labels

#add tip labels for pupil shape
tiplabels(col = "black", bg = "black",
          pch = sh_shape[pupils.eye$Final_Shape], #shape of labels
          cex = 0.9,
          offset = 15) #size of labels

#add tip labels for eye size
tiplabels(cex = 0.17*pupils.eye$eye_av,
          pch = 19, #shape of labels
          col = "mediumpurple",
          offset = 35) 


#add legend for pupil constriction
legend(x = 40, y = 209,
       legend = c("Horizontally elongated", "Non-elongated", "Vertically elongated"), 
       col = col_constrict,
       pch = sh_constrict, #shape of labels
       cex = 0.7, 
       box.lty = 0, 
       title = "Adult pupil", 
       title.adj = 0)

#add legend for pupil shape
legend(x = 40, y = 206, legend = c("Almond", "Circle", "Diamond", "Slit", "Inverted tear","Inverted triangle"),
       col = "black", pt.bg = "black",
       pch = sh_shape, #shape of labels
       cex = 0.7, 
       box.lty = 0, 
       title = "", 
       title.adj = 0)


#finish pdf export
#dev.off()
```

Alternatively, we can plot eye size with a bar plot in ggtree and color by pupil constriction. 

```{r fig.height = 15, fig.width = 4}

# Prep data and phylogeny-----

#subset data for eye diameter and constriction
eye.bars <- pupils.eye %>%
  mutate(tiplab = ASW_names,
         fam = Family,
         abseye = eye_av, 
         pupil = Final_Constriction, 
         pupil_bi = binary_constriction) %>%
  select(tiplab, fam, abseye, pupil, pupil_bi)

# set row names in dataset to match the tip labels in the tree
row.names(eye.bars) <- eye.bars$tiplab

#check that phylogeny and data match exactly
name.check(eye.bars, eye.tree)

#ladderize tree
eye.tree <- ladderize(eye.tree)

#resort trait dataset to the order of tree tip labels
eye.bars <- eye.bars[eye.tree$tip.label, ] 

#make trait vector for absolute eye size
aveye <- as.vector(eye.bars$abseye) 

#add tip label names to vector
names(aveye) <- eye.bars$tiplab 

#make trait vector of pupil constrictions
pups <- as.vector(eye.bars$pupil) 

#make vector of colors corresponding to phylogeny tips
tipcols.pup <- unname(col_constrict[pups]) 

# Make plot------

#export pdf 
#pdf(file = "../Outputs/Figures/eyesize_phylo.pdf", height=20, width=7)

#call plot with phylo, tip labels, and absolute eye diameters
plotTree.wBars(eye.tree, aveye, 
               scale = 6, 
               width = 1,
               tip.labels = FALSE, 
               offset = 0.4,
               col = tipcols.pup)

#add labels for family
#tiplabels(eye.bars$fam, cex = 0.7, adj = 1) 

#add legend for habitat states
legend(x  ="bottomleft", legend = c("Horizontally elongated", "Non-elongated", "Vertically elongated"), pch = 22, pt.cex= 2, pt.bg = col_constrict, cex = 1, bty = "n", horiz = F)

#dev.off()
```

## Test for correlation 

Here, we use a PGLS in caper to test whether there is a correlation between eye size (absolute eye diameter) and pupil shape (elongated vs. non-elongated) while accounting for evolutionary relationships. 

First we put our data and tree into a matched comparative object for caper. 

```{r}

#check that tree tip labels match data subset
name.check(eye.tree, pupils.eye)

#make sure order is same in tree and data
pupils.eye <- pupils.eye[eye.tree$tip.label,]

#use caper function to combine phylogeny and data into one object 
#(this function also matches species names in tree and dataset)
pupil_eye.comp <- comparative.data(phy = eye.tree, data = pupils.eye, 
                             names.col = ASW_names, 
                             vcv = TRUE, 
                             na.omit = FALSE, 
                             warn.dropped = TRUE)

#check for dropped tips or dropped species
pupil_eye.comp$dropped$tips #phylogeny
pupil_eye.comp$dropped$unmatched.rows #dataset
```

Next we can fit the PGLS model for pupil constriction vs. eye diameter. 

```{r}
#elongated pupils vs. eye diameter
pgls_pupil.eye <- pgls(eye_av ~ binary_constriction, 
               data = pupil_eye.comp,
               lambda = "ML", 
               param.CI = 0.95)
```

We need to check that model assumptions are being met. 

```{r}
#evaluate model assumptions
par(mfrow = c(2,2)) #makes your plot window into 2x2 panels
plot(pgls_pupil.eye) #plot the linear model
par(mfrow = c(1,1)) #set plotting window back to one plot
```

These look ok. Next we look at the model parameter estimates. 

```{r}
#main effects
anova(pgls_pupil.eye)

#parameter estimates
summary(pgls_pupil.eye)
```

Among species we have eye size data for from the ProcB paper (n = 207), there is a significant association between pupil constriction axis and eye diameter. Species with elongated pupils have significantly larger eyes than species with non-elongated pupils.

We can make boxplots comparing eye size across pupil types. 

```{r, fig.height=4, fig.width=10}


#shapes for boxplots
sh_constrict2 <- c("horizontal" = 22,
                   "symmetrical" = 21,
                   "vertical" = 25)

# boxplot of eye size across pupil constrictions
boxplot_eyesize <- ggplot(data = pupils.eye, 
                       aes(y = binary_constriction, x = eye_av)) + 
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha = 0.9) + #controls boxes  
 geom_jitter(aes(fill = Final_Constriction, shape = Final_Constriction), size = 3, alpha = 0.7, position = position_jitter(0.15)) +
  scale_y_discrete(limits = c("elongated", "symmetrical"),
                   labels = c("elongated", "non-elongated")) +
  scale_fill_manual(values = col_constrict,
                    name = "",
                    breaks = c("symmetrical", "horizontal", "vertical"),
                   labels = c("Non-elongated", "Horizontally elongated", "Vertically elongated")) +
  stat_summary(fun.x = mean, colour = "black", geom = "point", shape = 18, size = 4, show_guide = FALSE) + #controls what stats shown
  scale_shape_manual(values = sh_constrict2,
                     name = "",
                     breaks = c("symmetrical","horizontal", "vertical"),
                     labels = c("Non-elongated", "Horizontally elongated", "Vertically elongated")) +
  theme(text = element_text(size=16), panel.background = element_blank(), axis.line = element_line(colour = "black"), legend.key = element_rect(fill = NA)) + #controls background +
  ylab("Pupil shape") +
  xlab("Eye diameter (mm)")

boxplot_eyesize

#export pdf 
pdf(file = "../Outputs/Figures/eyesize_boxplot.pdf", height=4, width=11)
boxplot_eyesize
dev.off()


```

Let's look at this with 3 states

```{r}
# boxplot of eye size across pupil constrictions
boxplot_eyesize2 <- ggplot(data = pupils.eye, 
                       aes(x = eye_av, y = Final_Constriction)) + 
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha = 0.9) + #controls boxes  
 geom_jitter(size = 2, alpha = 0.7, position = position_jitter(0.15)) +
  stat_summary(fun.y = mean, colour = "black", geom = "point", shape = 18, size = 3, show_guide = FALSE) + #controls what stats shown
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black"), legend.key = element_rect(fill = NA)) + #controls background +
  xlab("eye diameter (mm)") +
  ylab("")

boxplot_eyesize2

```

